---
title: "NVTA VisionEval Scenario Results"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---


<!-- Add the following line manually to the rendered HTML document so that IE does not block the javascript elements: -->
<!-- saved from url=(0014)about:internet --> 


# Metropolitan Data Visualization 

The Volpe team started by looking at the daily VMT for all scenarios. This is done by combining seven DVMT outputs by location and type of mode. Percent change is calculated relative to the base scenario value.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)

# This sections sets the location of the VisionEval installation for each of the Volpe team members.
# The result is that the object `ve.runtime` is the full path to where VisionEval is installed, and the results from the NVTA scenarios are located. 
ve.runtime <- ifelse(grepl('Flynn', normalizePath('~/')), 
                     file.path(dirname('~/'), 'Desktop/VE_4-1-0/'),
                      ifelse(grepl('lohmar', normalizePath('~/')), 
                       'C:/VisionEval',
                         ifelse(grepl('englin', normalizePath('~/')), 
                         file.path('C:/Users/eric.englin/Desktop/VisionEval/4_04_v3/'),
                          NA)))

# Make a directory to save JPEG outputs as well
fig_save = file.path(ve.runtime, 'models', 'NVTA_Figures')
if(!dir.exists(fig_save)) { dir.create(fig_save) }

knitr::opts_knit$set(root.dir = ve.runtime)

library(tidyverse)
library(plotly) #install.packages('plotly')
library(DT) # install.packages('DT')

# Read in scenario descriptions
scenario_desc <- readxl::read_excel(file.path(ve.runtime,
                                              'models',
                                              'Scenarios_Overview.xlsx'))

# clean up trailing whitespace
scenario_values <- scenario_desc %>%
  mutate(`Model Run Name` = gsub('\\s+', '', `Model Run Name`))

scenario_values <- scenario_values %>%
  mutate(ModelNameLabel = ifelse(scenario_values$Level == 'NA' |
        str_detect(scenario_values$Folder,'C0\\d'), scenario_values$`Model Run Name`,
        paste0(scenario_values$`Model Run Name`, scenario_values$Level)))

scenario_values <- scenario_values %>%
  mutate(`ModelNameLabel` = gsub('\\s+', '', `ModelNameLabel`))

```




## DVMT Plots

### Percent change 

Across the study area, the largest DVMT changes in 2045 were observed for scenario K2 - Increasing Telecommuting, using the 'relative employment' input file in VisionEval to represent telecommuting. Increases in DVMT were observed for the 'larger families' scenario. Increases in DVMT were observed for the higher population growth scenario, which is as expected. 

```{r dvmt_rel}

marea_results <- read.csv(file.path(ve.runtime, 'models', 'Scenario_Metrics_Marea.csv'))

marea_results <- marea_results %>% 
  mutate(`modelName` = gsub('\\s+', '', `modelName`)) %>% 
  mutate(model_run_no = str_extract(modelName, '[A-Z]\\d{1,3}$')) %>% 
  left_join(scenario_values, by = c("modelName" = "ModelNameLabel")) 


marea_results <- marea_results %>%   
  mutate(modelName = ifelse(modelName == 'VERSPM_NVTA', 
                            '0 - Base',
                            paste(model_run_no, `Short Name`, sep = ' - ')))

#info for regional dvmt graphic

dvmt <- marea_results %>%
  select(modelName, contains('Dvmt')) %>%
  rowwise() %>%
  mutate(regionalDVMT = sum(c_across(contains('Dvmt'))))

dvmtbase = as.numeric(dvmt %>% filter(modelName == '0 - Base') %>% select(regionalDVMT))
#dvmtbase = as.numeric(dvmt %>% filter(modelName == 'VERSPM_NVTA') %>% select(regionalDVMT))

dvmt <- dvmt %>%
  mutate(regDVMTAbsDif = regionalDVMT - dvmtbase,
         regDVMTPerChange = (regDVMTAbsDif/dvmtbase)*100,
         color = ifelse(regDVMTPerChange > 0, "pos","neg"),
         pct_text = paste('Scenario:', modelName, '\n', round(regDVMTPerChange, 2), '%'),
         mi_text = paste('Scenario:', modelName, '\n', format(round(regDVMTAbsDif, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'mi/day'))

gp1 <- ggplot(dvmt, aes(x = modelName, y = regDVMTPerChange, fill = color,
                        text = pct_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change Regional DVMT", x = "Scenario", y = "Percent Change") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# Save the output figure to disk for inclusion in a manuscript
ggsave(file.path(fig_save, 'DVMT_Rel_Change.jpeg'))

# And plot dynamically for viewing.
ggplotly(gp1, tooltip = 'text')
```

### Absolute change

```{r dvmt_abs}
gp2 <- ggplot(dvmt, aes(x = modelName, y = regDVMTAbsDif, fill = color,
                        text = mi_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Absolute Change Regional DVMT (mi/day)", x = "Scenario", y = "Absolute Change")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'DVMT_Abs_Change.jpeg'))

ggplotly(gp2, tooltip = 'text')
```

## Energy Usage plots

### Truck energy usage relative change

As with the DVMT values, just looking at energy usage for heavy trucks, the largest decrease was for scenario H2 - increasing hybrid and battery electric trucks. Substantial decreases were also observed for K2 - Increasing Telecommuting. This is worth noting because we would not necessarily expect heavy truck trips to be altered so much by telecommuting, so this deserves more investigation. Increases in heavy truck energy use were observed for the 'L3 - Larger Families' scenario.  

The combination scenarios C02 and C05, which include the larger families as well as changes to employment, population, and fuels tax, had similar energy usage changes.

```{r truck_energy_rel}

#info for commercial heavy truck gge (sum columns k-m)
# use `matches()` to use regular expressions first to find variable names that start with 
# ComSvc or HvyTrk, and then a second one to find only those which also have GGE at the end
gge <- marea_results %>%
  select(modelName, matches('^(ComSvc|HvyTrk)')) %>%
  select(modelName, matches('GGE$')) %>%
  rowwise() %>%
  mutate(ComHvyTruckGGE = sum(c_across(contains('GGE'))))

ggebase = as.numeric(gge %>% filter(modelName == '0 - Base') %>% select(ComHvyTruckGGE))
#ggebase = as.numeric(gge %>% filter(modelName == 'VERSPM_NVTA') %>% select(ComHvyTruckGGE))

gge <- gge %>%
  mutate(ComHvyTruckGGEAbsDif = ComHvyTruckGGE - ggebase,
         ComHvyTruckGGEPerChange = (ComHvyTruckGGEAbsDif/ggebase)*100,
         sign = ifelse(ComHvyTruckGGEPerChange > 0, "pos","neg"),
         pct_text = paste('Scenario:', modelName, '\n', round(ComHvyTruckGGEPerChange, 2), '%'),
         mi_text = paste('Scenario:', modelName, '\n', format(round(ComHvyTruckGGEAbsDif, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'Gasoline gallone equiv.'))

# View(gge)

gp3 <- ggplot(gge, aes(x = modelName, y = ComHvyTruckGGEPerChange, fill = sign,
                       text = pct_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change Commercial Service and  Heavy Truck GGE", x = "Scenario", y = "Percent Change")+
  theme(axis.text.x = element_text(angle = 45))

ggsave(file.path(fig_save, 'Truck_GGE_Pct_Change.jpeg'))

ggplotly(gp3, tooltip = 'text')

```

## Metropolitan area data table


```{r marea_table}
marea_table <- marea_results %>% select(-names(marea_results)[names(marea_results) %in% names(scenario_values)])

val_length <- sapply(marea_table, FUN = function(x) ifelse(is.numeric(x), max(nchar(as.character(round(x, 0)))),
                                                           max(nchar(x))))



datatable(marea_table,
          rownames = F,
          filter = 'top',
          extensions = c('Buttons','ColReorder', 'Select'),
 options = list(
   pageLength = 10,
   dom = 'BSfrtip',
   colReorder = TRUE,
   select = TRUE,
   search = list(regex = TRUE), # add regex searching to all search boxes
   buttons = c('csv', 'excel', 'colvis'),
   autoWidth = TRUE,
   scrollX = TRUE, scrollY = 600)
  ,
 caption = "Complete metropolitan area results for NVTA scenarios." ) %>%
  formatRound(which(val_length >= 5 & !grepl('model', names(val_length))), 0) %>%
  formatRound(which(val_length < 5 & !grepl('model', names(val_length))), 2) 

```

# Household Data Visualization

In addition to examining the outputs which are reported at the entire metropolitan area level, we also investigated the outputs rolled up across all households. 

## Energy consumption - gasoline equivalents

```{r household_prep, cache = F}
hhfile <- readr::read_csv(file.path(ve.runtime, 'models', 'Scenario_Metrics_Hh.csv'))

# Join in the scenario names
hhfile <- hhfile %>%
  mutate(`modelName` = gsub('\\s+', '', `modelName`)) %>% 
  mutate(model_run_no = str_extract(modelName, '[A-Z]\\d{1,3}$')) %>% 
  left_join(scenario_values, by = c("modelName" = "ModelNameLabel")) 

hhfile <- hhfile %>%
  mutate(modelName = ifelse(modelName == 'VERSPM_NVTA', 
                            '0 - Base',
                            paste(model_run_no, `Short Name`, sep = ' - ')))



# Calculate mode split
hhfile <- hhfile %>%
  rowwise() %>%
  mutate(total_trips = sum(WalkTrips, BikeTrips, TransitTrips, VehicleTrips),
         pct_Walk = WalkTrips / total_trips,
         pct_Bike = BikeTrips / total_trips,
         pct_Transit = TransitTrips / total_trips,
         pct_Vehicle = VehicleTrips / total_trips)

scenario_values <- hhfile %>%
  filter(modelName != '0 - Base') %>% # exclude the base scenario
  group_by(modelName) %>%
  summarize(sum_Hh_GGE = sum(DailyGGE),
            sum_Hh_kWh = sum(DailyKWH),
            sum_Hh_CO2e = sum(DailyCO2e),
            median_Income = median(Income),
            median_OwnCost = median(OwnCost, na.rm = T),
            median_pct_Walk = median(pct_Walk),
            median_pct_Bike = median(pct_Bike),
            median_pct_Transit = median(pct_Transit),
            median_pct_Vehicle = median(pct_Vehicle)
            )

base_values <- hhfile %>%
  filter(modelName == "0 - Base") %>%
  group_by(modelName) %>%
  summarize(base_Hh_GGE = sum(DailyGGE),
            base_Hh_kWh = sum(DailyKWH),
            base_Hh_CO2e = sum(DailyCO2e),
            median_Income = median(Income),
            median_OwnCost = median(OwnCost, na.rm = T),
            median_pct_Walk = median(pct_Walk),
            median_pct_Bike = median(pct_Bike),
            median_pct_Transit = median(pct_Transit),
            median_pct_Vehicle = median(pct_Vehicle)
            )

# We use base_values$... to compare the scenario values with the corresponding base values data frame

scenario_values <- scenario_values %>%
  mutate(GGEPerChange = ((sum_Hh_GGE - base_values$base_Hh_GGE) / base_values$base_Hh_GGE) * 100,
         KWHPerChange = ((sum_Hh_kWh - base_values$base_Hh_kWh) / base_values$base_Hh_kWh) * 100,
         CO2PerChange = ((sum_Hh_CO2e - base_values$base_Hh_CO2e) / base_values$base_Hh_CO2e) * 100,
         gge_color= ifelse(GGEPerChange > 0, "pos","neg"),
         gge_pct_text = paste('Scenario:', modelName, '\n', round(GGEPerChange, 2), '%'),
         gge_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_GGE, 2),
                                                               scientific = F, 
                                                               big.mark = ','), 'Gasoline gallone equiv.'),
         kwh_color= ifelse(KWHPerChange > 0, "pos","neg"),
         kwh_pct_text = paste('Scenario:', modelName, '\n', round(KWHPerChange, 2), '%'),
         kwh_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_kWh, 2),
                                                               scientific = F, 
                                                               big.mark = ','), 'kWh'),
         co2_color= ifelse(CO2PerChange > 0, "pos","neg"),
         co2_pct_text = paste('Scenario:', modelName, '\n', round(CO2PerChange, 2), '%'),
         co2_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_CO2e, 2),
                                                               scientific = F, 
                                                               big.mark = ','), 'g CO[2] equivalent'))


gp4 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_GGE,
                                   text = gge_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional daily gasoline gallon equivalents", x = "Scenario", y = "GGE/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_GGE_Total.jpeg'))

ggplotly(gp4, tooltip = 'text')
```

```{r householde_ggepct}
gp5 <- ggplot(scenario_values, aes(x = modelName, y = GGEPerChange,
                                   text = gge_pct_text, fill = gge_color)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Regional daily gasoline gallon equivalents % Change", x = "Scenario", y = "GGE/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_GGE_Percent_Change.jpeg'))

ggplotly(gp5, tooltip = 'text')
```

### Energy consumption - electricity 

```{r household_kwh}
gp6 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_kWh,
                            text = kwh_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional KWH", x = "Scenario", y = "kWh/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_kWh_Total.jpeg'))


ggplotly(gp6, tooltip = 'text')
```

Again, scenario K2 - Increasing telecommuting had by far the biggest reduction in electricity use. 

The combination scenarios C01 and CO4 both include the increased telecommuting scenario, along with electrification and increased funding to transit.

```{r household_kwh_pct}
gp7 <- ggplot(scenario_values, aes(x = modelName, y = KWHPerChange,
                            text = kwh_pct_text, fill = kwh_color)) + 
  geom_bar(stat="identity") + theme_minimal() +
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Regional KWH Percent Change", x = "Scenario", y = "kWh/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



ggsave(file.path(fig_save, 'Regional_kWh_Percent_Change.jpeg'))

ggplotly(gp7, tooltip = 'text')
```

### Emissions

```{r household_emissions}
gp8 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_CO2e,
                            text = co2_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional CO[2]e", x = "Model Name", y ="g CO[2] equivalent / Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggsave(file.path(fig_save, 'Regional_CO2e_Total.jpeg'))

ggplotly(gp8, tooltip = 'text')
```

The largest changes in emissions, summed up from the household CO[2]e values, was under the scenario with increasing electrification (2 different levels of increase tested), with the very substantial reductions as well observed under the K2 - increasing telecommuting scenario. 

The best combination (C01) includes G2 + H2 + K2, meaning increased household vehicle electrification, increase in hybrid heavy trucks, along with increased telecommuting.

```{r household_emissions_pct}
gp9 <- ggplot(scenario_values, aes(x = modelName, y = CO2PerChange,
                            text = co2_pct_text, fill = co2_color)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional CO[2]e Percent Change", x = "Model Name", y ="g CO[2] equivalent / Day")+
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggsave(file.path(fig_save, 'Regional_CO2e_Percent_Change.jpeg'))

ggplotly(gp9, tooltip = 'text')


```

### Mode Choice

The biggest reduction in vehicle usage (as measured by the median percent of trips at the household level for each of vehicle, bicycle, transit, and walking) was for the increased telecommuting scenario, along with the combination scenarios which included this telecommuting shift. 

```{r mode_choice}
# ranking by  % vehicle trips

mode_choice <- scenario_values %>%
  select(modelName, contains('median_pct')) %>%
  pivot_longer(cols = contains('median_pct'),
               names_to = 'Mode',
               values_to = 'median_pct') %>%
  mutate(Mode = gsub('median_pct_', '', Mode))



gp10 <- ggplot(mode_choice, aes(x = modelName, y = median_pct, fill = Mode)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Percent of household trips by mode", x = "Model Name", y ="Median % of trips by each mode") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggsave(file.path(fig_save, 'Mode_Choice.jpeg'))

ggplotly(gp10)




```


### Ownership Costs

The greatest reduction in the costs to households of vehicle ownership was from the aging in place scenario.

```{r own_cost}
gp11 <- ggplot(scenario_values, aes(x = reorder(modelName, median_OwnCost), y = median_OwnCost)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Median vehicle ownership costs", x = "Model Name", y ="$ / household") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggsave(file.path(fig_save, 'Ownership_costs.jpeg'))

ggplotly(gp11)
```

## Household level data table


```{r hh_table}
scenario_table <- scenario_values %>% select(names(scenario_values)[!grepl('_text', names(scenario_values)) &
                                                 !grepl('_color', names(scenario_values))])

val_length <- sapply(scenario_table, FUN = function(x) ifelse(is.numeric(x), max(nchar(as.character(round(x, 0)))),
                                                           max(nchar(x))))

datatable(scenario_table,
          rownames = F,
          filter = 'top',
          extensions = c('Buttons','ColReorder', 'Select'),
 options = list(
   pageLength = 10,
   dom = 'BSfrtip',
   colReorder = TRUE,
   select = TRUE,
   search = list(regex = TRUE), # add regex searching to all search boxes
   buttons = c('csv', 'excel', 'colvis'),
   autoWidth = TRUE,
   scrollX = TRUE, scrollY = 600)
  ,
 caption = "Summary of household level results for NVTA scenarios." ) %>%
  formatRound(which(val_length >= 5 & !grepl('model', names(val_length))), 0) %>%
  formatRound(which(val_length < 5 & !grepl('model', names(val_length))), 2)  

```

# Validation 

The Volpe team also realized the need to validate the VisionEval output. 

#### MWCOG DVMT

The VDOT team sent along daily VMT values in 2021 and 2045 from the MWCOG MOE Base model. These results include household car VMT, business car VMT, and car travel from other locations (such as motorists who live outside the area and are passing through it.

Here are the raw results from the MWCOG base model in 2045:

       Jurisdiction  Freeway     Expressway  Major Art.  Minor Art.   Collector     Ramp       All
-------------------- ----------- ----------- ----------- ----------- ----------- ----------- -----------
 Fairfax County      12,517,806   5,910,760   6,219,736   1,419,677   1,256,563     488,882  27,813,424
 City of Fairfax              0     284,867      99,406       9,144           0           0     393,417
 Falls Church                 0     106,172      40,693       4,605           0           0     151,470
 
 
CROSSTAB ROW=FLAG COL=FTYPE VAR=_24VMTTRK
                                                                                                       
       Jurisdiction  Freeway     Expressway  Major Art.  Minor Art.   Collector     Ramp       All
-------------------- ----------- ----------- ----------- ----------- ----------- ----------- -----------
Fairfax County        3,071,425     988,206     919,784     212,966     331,425      90,603   5,614,408
City of Fairfax               0      37,640      12,957         774           0           0      51,371
Falls Church                  0      18,944       5,575         425           0           0      24,944




When you add all of the numbers up, you get 34,049,034 daily VMTs. This is about 4,000,000 higher than the VisionEval dvmt values. The chart below shows the percent differences with each of our scenarios. 

```{r dvmt_plot_validation}

ffx_county_car_2045 = 27813424
ffx_city_car_2045 = 393417
fall_church_car_2045 = 151470
ffx_county_truck_2045 = 5614408
ffx_city_truck_2045 = 51371
fall_church_truck_2045 = 24944

MWCOG_model_dvmt = ffx_county_car_2045+ffx_city_car_2045+fall_church_car_2045+ffx_county_truck_2045+ffx_city_truck_2045+fall_church_truck_2045

#trip_generation_summary_motorized_dvmt = 3649551
#trip_generation_summary_non_motorized_dvmt = 317247

dvmt$regDVMTAbsDif = dvmt$regionalDVMT - MWCOG_model_dvmt
dvmt$regDVMTPerChange = (dvmt$regDVMTAbsDif/MWCOG_model_dvmt)*100
dvmt$color <- ifelse(dvmt$regDVMTPerChange < 0, "neg","pos")

dvmtplot1 <- data.frame(Model = dvmt$modelName,PercentChange = dvmt$regDVMTPerChange, 
                        color = dvmt$color,
                        vmt_text = paste('Scenario:', dvmt$modelName, '\n', round(dvmt$regDVMTPerChange, 2), '%'))


gp_v <- ggplot(dvmtplot1, aes(x=Model, y=PercentChange, fill = color, text = vmt_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change VisionEval DVMT and MWCOG MOES Base Model for 2045", x = "Model Name", y = "Percent Change")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'VMT_Validation.jpeg'))

ggplotly(gp_v, tooltip = 'text')



```

```{r valid_table}

library(DT)

datatable(dvmt %>% select(-regDVMTAbsDif, -color, -pct_text, -mi_text) %>%
            relocate(regionalDVMT, .after = modelName),
           rownames = F,
          filter = 'top',
          extensions = c('Buttons'),
          options = c(
            pageLength = 10,
            dom = 'BSfrtip',
            colReorder = TRUE,
            select = TRUE,
            search = list(regex = TRUE), # add regex searching to all search boxes
            buttons = list('copy', 'print',
                           list(
                             extend = 'collection',
                             buttons = c('csv', 'excel', 'pdf'),
                             text = 'Download')),
            autoWidth = TRUE,
            scrollX = TRUE, scrollY = 600),
           caption = "VMT values from NVTA scenarios" ) %>%
  formatRound(2:9, 0) %>%
  formatStyle('regionalDVMT', 
              backgroundColor = 'dodgerblue')

```

```{r save}
save(file = 'Final_NVTA_results.RData', 
     list = c('marea_results',
              'scenario_desc',
              'scenario_values',
              'hhfile'))

# to confirm the total CO2 values from household file:
# co2compile <- hh_compiled %>% group_by(modelName) %>% summarize(sum_CO2 = sum(DailyCO2e))
# View(co2compiled)

```